\sysfull

### 2017-11-06 paper/intro.tex
Revised text in intro after Sinan's comments; remarks about Figure 1

There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl}. New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now). However, they have their own limitations that need to be addressed---the core one being inability to adapt to changing energy conditions.

\textbf{Static Task-based Model and its Inefficiency:} The latest two {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks. Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures. The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent. A key challenge is that the length of a software task's execution is \emph{limited by the fixed total amount of energy} that a device can buffer in hardware. A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict. To illustrate this, we refer to Figure~\ref{fig:1} showing the execution time of two applications (the same application X, but divided into X and Y tasks as in~\cite{chain}) running on RFID antenna-powered Computational RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna distances (near---stable energy supply/far---high energy intermittency). At short distance: execution will takes too long caused by runtime cost of marshalling excessively small tasks. At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer. This calls for at-runtime task division of any transiently-powered application.

\textbf{Data Inconsistency of Virtual Task:} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task. This however induces a problem of \emph{virtual task write-after-read (WAR) dependency violation}. Consider the following example of Task 1: {\tt{\{y=x+1;\}}}, Task 2: {\tt{\{z=y+1;\}}}, and Task 3 {\tt{\{x=z+1;\}}} being executed consecutively. All tasks have no WAR dependency and are consistent. Now, suppose three tasks have been virtualized into one, namely Task 4: {\tt{\{y=x+1; z=y+1; x=z+1;\}}} which has inherent WAR dependency (variable {\tt{x}}).

\textbf{Trade-off with Virtual Memory Manipulation:} WAR dependencies cannot be complied at runtime for merged (virtual) tasks. Therefore virtual task execution model needs to manipulate merged tasks in-memory to guarantee data consistency. Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}. Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}). How to connect this trade-off with task virtualization remains an open question.

@@

To address this question, in this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}. To accomplish this, \sys introduces two new capabilities:

\begin{itemize}
	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism dynamically executes multiple tasks as a single task. A programmer or compiler can specify small tasks that will execute on a device with a small energy buffer. Coalescing tasks allows running the same tasks on a device with a larger energy buffer, avoiding run time overheads associated with ending one task and beginning another.
	%
	\item {\bf Volatile memory virtualization on transient power:} Colascing through task virtualization requires memory virtualization. \sys's uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution. When a task ends, \sys commits dirty pages to FRAM using DMA block copies, ensuring task atomicity and memory consistency. To the best of our knowledge this is the first time when memory virtualization has been demonstrated on a transient power budget.
\end{itemize}

>>>>

There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl}. New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now). However, they have their own limitations that need to be addressed---the core one being \emph{inability to adapt} to changing energy conditions.

\textbf{The Need for Adaptive Task Division:} Recently proposed {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks. Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures. The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent. A key challenge is that the length of a software task's execution is \emph{limited by the fixed total amount of energy} that a device can buffer in hardware. A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict. To illustrate this, we refer to Figure~\ref{fig:1} showing the execution time of two applications (the same application X, but divided into X and Y tasks as in~\cite{chain}) running on RFID antenna-powered Computational RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna distances (near---stable energy supply/far---high energy intermittency). At short distance: execution will takes too long caused by runtime cost of marshalling excessively small tasks. At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer. This calls for at-runtime adaptive task division of any transiently-powered application---namely \emph{task virtualization}.

\begin{figure}
	\includegraphics[width=.5\columnwidth]{presentation/images/alpaca-war.pdf}
	\caption{\label{fig:virtualization}Adaptive task division at run-time might lead to write-after-read dependencies (variable \ttfamily{x}) that do not exist with static task division.\todo{Box with Task 1,2,3 must be the same size as merged task; Merged task must be named as Task 4; x must be marked in text as having WAR; "Task X" name should be next to the task box; "Window" - small letter}{Sinan}}
\end{figure}

\textbf{The Challenge of Task Virtualization:} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task. This however might introduce \emph{write-after-read (WAR)} dependencies on the variables in non-volatile memory at \emph{run-time} that can break the atomicity of the \emph{virtual task}. Consider the example depicted in Fig.~\ref{fig:virtualization}: Tasks 1, 2 and 3 being executed consecutively. All tasks in this example are atomic since they do not have WAR dependency on the persistent variables they are accessing, e.g. \emph{x} is only read and \emph{y} is only written within Task 1. Now, suppose three tasks have been virtualized into a single one at run-time, namely Task 4: since \emph{x} is now first \emph{read} and then \emph{written}, a WAR dependency on \emph{x} is introduced dynamically at run-time---unfortunately Task 4 is \emph{no more atomic} since its re-execution will not always produce the same results.

\textbf{Guaranteeing Atomicity of Virtual Tasks:} Although WAR dependencies introduced by the static task division can be eliminated at compile time, e.g.~\cite{alpaca}, those introduced dynamically at run-time, due to task virtualization, cannot. Therefore, we require a \emph{new execution model} that keeps each virtual task atomic by committing the modified persistent variables at the boundary defined dynamically at run-time---keeping the non-volatile memory unmodified upon a power-interrupt and preserving its consistency.

%Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}. Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}). How to connect this trade-off with task virtualization remains an open question.

@@

To address this question, we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}. To accomplish this, \sys introduces two new capabilities:

\begin{itemize}
	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism dynamically executes multiple tasks as a single task. A programmer or compiler can specify small tasks that will execute on a device with a small energy buffer. Coalescing tasks allows running the same tasks on a device with a larger energy buffer, avoiding run time overheads associated with ending one task and beginning another.
	%
	\item {\bf Volatile memory virtualization with transiently-powered device:} Colascing through task virtualization requires memory virtualization. \sys's uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution. When a task ends, \sys commits dirty pages to FRAM using DMA block copies, ensuring task atomicity and memory consistency. To the best of our knowledge this is the first time when memory virtualization has been demonstrated on a transient power budget.
\end{itemize}
